### 2022-05-05 README.Rmd
added readme

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Comparing corrosion control treatments using a robust Bayesian generalized additive model

This repository contains the R code and other materials for the paper "Comparing corrosion control treatments using a robust Bayesian generalized additive model".