### 2022-05-31 README.Rmd
readme update

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Comparing corrosion control treatments for drinking water using a robust Bayesian generalized additive model

This repository contains the R code and other materials for the paper "Comparing corrosion control treatments for drinking water using a robust Bayesian generalized additive model", currently available as a [preprint](http://doi.org/10.26434/chemrxiv-2022-b3f72-v3). Bespoke functions used to fit the models are available as a separate package, [`bgamcar1`](https://github.com/bentrueman/bgamcar1).